---
title: "Have ICT changed the game for mobility and migration?"
subtitle: "Working Paper. Please don't cite"
author: "Ayra Rowena Reyla and Ana Cecilia Montes"
date: "11 November 2015"
output: pdf_document
bibliography: referencesreylamontes.bib

highlight: zenburn
theme: united
toc: yes
toc_depth: 3
---
\newpage
---
 1 Introduction

##5.2 Emperical Strategy

#Emperical Methods


#Descriptive Statistics

```{r setup, include=FALSE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#install.packages('WDI')
#install.packages('tidyr')
#install.packages('rio')
#install.packages('countrycode')
#install.packages("RJOSONIO")  
#install.packages ("ggplot2")
#install.packages("rworldmap")
#install.packages("sp")
#install.packages("joinCountryData2Map")
#install.packages("plm")
#install.packages("Formula")
#install.packages("pglm")

# Loading packages
library("maptools")
library("countrycode")
library("RJSONIO")
library("WDI")
library("tidyr")
library("rio")
library("ggplot2")
library("rworldmap")
library("sp")
library('Formula')
library('plm')
library('pglm')

# Setting directory
setwd('/Users/AnaCe/Desktop/Final_ProjectMontesReyla')
#setwd('/Users/ayrarowenareyla/Desktop/The Hertie School of Governance/Collaborative Social Sciences/Final_ProjectMontesReyla')

#loading data set
Merged <- read.csv(file="MontesandReyla")
Merged2 <- read.csv(file="MontesandReylaFinalVars")

#subdataframes
merged00 <- read.csv(file="merged00")
merged10 <- read.csv(file="merged10")
merged13 <- read.csv(file="merged13")

```

## Dependent Variables

## Summary

### Patterns of Emmigration

#Independent Variables

#Results


#Table 1 Panel Regression Cellphone users
<center>
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results="asis", header=FALSE, font.size="tiny"}
#Panel regression by years and interaction
#Cellphones and Within

Within_11<- plm(logemigrationpercap ~ CellphoneUsers*logGDPPerCapita +  FertilityRate + PoliticalStability + employmentprob, data = Merged, index = c("country", "year"), model = "within")

Within_12 <- plm(logemigrationpercap ~ CellphoneUsers + logGDPPerCapita +  FertilityRate + PoliticalStability + employmentprob, data = Merged, index = c("country", "year"), model = "within")

Within_13<- plm(logemigrationpercap ~ CellphoneUsers+  FertilityRate + PoliticalStability + employmentprob, data = Merged, index = c("country", "year"), model = "within")

Within_14<- plm(logemigrationpercap ~ CellphoneUsers+  PoliticalStability + employmentprob, data = Merged, index = c("country", "year"), model = "within")

Within_15<-plm(logemigrationpercap ~ CellphoneUsers+ employmentprob, data = Merged, index = c("country", "year"), model = "within")

Within_16<-plm(logemigrationpercap ~ CellphoneUsers, data = Merged, index = c("country", "year"), model = "within")

#Table
labels1<-c('CellphoneUsers', 'logGDPpp-1', 'Fertility Rate', 'Political Stability', 'Employment prob', 'CellphoneUsersXlogGDPpp-1')

table1<-stargazer::stargazer(Within_16, Within_15, Within_14, Within_13, Within_12, Within_11, type = "latex", header = FALSE, title = "Panel Regression of emigration rate", digits = 4, single.row=FALSE, omit.stat = c("f", "ser"), dep.var.caption="Emigration rate per cap (log)", covariate.labels = labels1)

```

#Table 2 Panel Regression internet usrer
<center>
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results="asis", header=FALSE, font.size="tiny"}
#Panel regression by years and interaction
#Cellphones and Within

Within_21<- plm(logemigrationpercap ~ InternetUsers*logGDPPerCapita +  FertilityRate + PoliticalStability + employmentprob, data = Merged, index = c("country", "year"), model = "within")

Within_22 <- plm(logemigrationpercap ~ InternetUsers + logGDPPerCapita +  FertilityRate + PoliticalStability + employmentprob, data = Merged, index = c("country", "year"), model = "within")

Within_23<- plm(logemigrationpercap ~ InternetUsers+  FertilityRate + PoliticalStability + employmentprob, data = Merged, index = c("country", "year"), model = "within")

Within_24<- plm(logemigrationpercap ~ InternetUsers+  PoliticalStability + employmentprob, data = Merged, index = c("country", "year"), model = "within")

Within_25<-plm(logemigrationpercap ~ InternetUsers+ employmentprob, data = Merged, index = c("country", "year"), model = "within")

Within_26<-plm(logemigrationpercap ~ InternetUsers, data = Merged, index = c("country", "year"), model = "within")

#Table
labels2<-c('InternetUsers', 'logGDPpp-1', 'Fertility Rate', 'Political Stability', 'Employment prob', 'InternetUsersXlogGDPpp-1')

table2<-stargazer::stargazer(Within_26, Within_25, Within_24, Within_23, Within_22, Within_21, type = "latex", header = FALSE, title = "Panel Regression of emigration rate", digits = 4, single.row=FALSE, omit.stat = c("f", "ser"), dep.var.caption="Emigration rate per cap (log)", covariate.labels = labels2)

```


#Table 3. Yearly regressions
<center>
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results="asis", header=FALSE, font.size="tiny"}
#OLS Regression by years and interaction
#Cellphones
OLS00_1 <- lm(logemigrationpercap ~ CellphoneUsers*logGDPPerCapita.l + FertilityRate + PoliticalStability + employmentprob, data = merged00)
OLS10_1 <- lm(logemigrationpercap ~ CellphoneUsers*logGDPPerCapita.l + FertilityRate + PoliticalStability + employmentprob, data = merged10)

OLS13_1 <- lm(logemigrationpercap ~ CellphoneUsers*logGDPPerCapita.l + FertilityRate + PoliticalStability + employmentprob, data = merged13)

# Internet
OLS00_2 <- lm(logemigrationpercap ~ InternetUsers*logGDPPerCapita.l + FertilityRate + PoliticalStability + employmentprob, data = merged00)

OLS10_2 <- lm(logemigrationpercap ~ InternetUsers*logGDPPerCapita.l + FertilityRate + PoliticalStability + employmentprob, data = merged10)

OLS13_2 <- lm(logemigrationpercap ~ InternetUsers*logGDPPerCapita.l + FertilityRate + PoliticalStability + employmentprob, data = merged13)

#Table 3

labels3 <-c('CellphoneUsers', 'InternetUsers', 'logGDPpp-1', 'Fertility Rate', 'Political Stability', 'Employment prob','CellphoneUsersXlogGDPpp-1', 'InternetUsersXlogGDPpp-1')
table3 <-stargazer::stargazer(OLS00_1, OLS10_1, OLS13_1, OLS00_2, OLS10_2, OLS13_2, column.labels = c("2000", "2010", "2013", "2000", "2010", "2013"), type = "latex", header = FALSE, title = "OLS Regression of emigration stocks around the world 2000-2013", digits = 4, single.row=FALSE, omit.stat = c("f", "ser"), dep.var.caption="Emigration rate per cap (log)", covariate.labels = labels3)
```

#Limitations and Further Research

#**References**


# Appendix
Table X Panel Regression - All model- Cellphone Users

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results="asis", header=FALSE, font.size="tiny"}
#Panel regression by years and interaction
#Cellphones

pooling_1 <- plm(logemigrationpercap ~ CellphoneUsers*logGDPPerCapita +  FertilityRate + PoliticalStability + employmentprob, data = Merged, index = c("country", "year"), model = "pooling")

#Within
Within_11 <- plm(logemigrationpercap ~ CellphoneUsers*logGDPPerCapita +  FertilityRate + PoliticalStability + employmentprob, data = Merged, index = c("country", "year"), model = "within")

Between_11 <- plm(logemigrationpercap ~ CellphoneUsers*logGDPPerCapita +  FertilityRate + PoliticalStability + employmentprob, data = Merged, index = c("country", "year"), model = "between")

Random_11 <- plm(logemigrationpercap ~ CellphoneUsers*logGDPPerCapita + FertilityRate + PoliticalStability + employmentprob, , data = Merged, index = c("country", "year"), model = "random")

# Table XXPanel Regression - All model- Internet USers
labels1<-c('CellphoneUsers', 'logGDPpp-1', 'Fertility Rate', 'Political Stability', 'Employment prob', 'CellphoneUsersXlogGDPpp-1')

table10<-stargazer::stargazer(pooling_1, Within_11, Between_11, Random_11, column.labels = c("Pool OLS", "Within", "Between", "Random"), type = "latex", header = FALSE, title = "Panel Regression of emigration stocks around the world 2000-2013", digits = 4, single.row=FALSE, omit.stat = c("f", "ser"), dep.var.caption="Emigration rate per cap (log)", covariate.labels = labels1)

```

#Panel regression - All Models- Internet users
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results="asis", header=FALSE, font.size="tiny"}
# Pooling

pooling_21<- plm(logemigrationpercap ~ InternetUsers*logGDPPerCapita +  FertilityRate + PoliticalStability + employmentprob, data = Merged, index = c("country", "year"), model = "pooling")

#Within
Within_21 <- plm(logemigrationpercap ~ InternetUsers*logGDPPerCapita +  FertilityRate + PoliticalStability + employmentprob, data = Merged, index = c("country", "year"), model = "within")

Between_21 <- plm(logemigrationpercap ~ InternetUsers*logGDPPerCapita +  FertilityRate + PoliticalStability + employmentprob, data = Merged, index = c("country", "year"), model = "between")

Random_21 <- plm(logemigrationpercap ~ InternetUsers*logGDPPerCapita + FertilityRate + PoliticalStability + employmentprob, , data = Merged, index = c("country", "year"), model = "random")

#Table
labels2<-c('InternetUsers', 'logGDPpp-1', 'Fertility Rate', 'Political Stability', 'Employment prob', 'CellphoneUsersXlogGDPpp-1')

table10<-stargazer::stargazer(pooling_21, Within_21, Between_21, Random_21, column.labels = c("Pool OLS", "Within", "Between", "Random"), type = "latex", header = FALSE, title = "Panel Regression of emigration stocks around the world 2000-2013", digits = 4, single.row=FALSE, omit.stat = c("f", "ser"), dep.var.caption="Emigration rate per cap (log)", covariate.labels = labels2)

```